---
title: "Final Project Q3"
author: "Wenjing Xu"
date: "October 31, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To implement K-Means algorithm, we should do following steps.

1. Randomly choose K points (which is 3 here) as the initial centers.

2. For each observation, determine which cluster center is the cloest to it using Euclidean distance measure. Then, assign that observation to the corresponding cluster.  

3. Compute the mean location of the points in each cluster. Set this mean location as the new cluster center point. 

4. Repeat step 2 and 3, until there is no more change for clusters.

```{r}
## Import Library and Data
library('rattle')
data(wine)
X <- wine[-1]

## Run k-means algorithm
run_kmeans <- function(data,k){
  data <- as.matrix(data)
  num_points <- nrow(data)
  num_measurements <- ncol(data)
  start_mean <- data[sample(1:num_points, k, replace=F),]
  new_index <- rep(0,num_points)
  for(i in 1:5){
    distance <- matrix(0,num_points,k)
    for (j in 1:k){
      distance[,j] <- apply((data-matrix(start_mean[j,],nrow=num_points,ncol=num_measurements,byrow=T))^2,1,sum)
    }
    new_index <- apply(distance,1,which.min)
    for (n in 1:k){
      start_mean[j] <- apply(data[which(new_index == n),],2,mean)
    }
  }
  return (new_index)
}

run_kmeans(X,3)

```



